Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-02-07 paper/main.tex
Add TODOs
Use V for velocity (including U^2 -> V^2)

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared".
% TODO: If U-squared is "velocity squared" then  make sure that's consistent
% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
% TODO: Decide on better terms than "fitted" and "not fitted".
% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
% complexity vs. accuracy at the time of convergence.
% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

@@

\subsection*{Analytical models for comparison}
The analytical models used for comparison are also listed in
Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces
added to a collection of aerodynamic terms.  The primary aerodynamic
contribution in all quasi-steady models is a translational lift term
proportional to the squared velocity of each span-element of the wing that is
related to the wing's angle of attack and empirically determined coefficient of
lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal
horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when
out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
\end{equation}
where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

>>>>

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared". -->> Change to V^2
% everywhere (see Whitney Wood paper just below equation 2.18).
% TODO: If U-squared is "velocity squared" then  make sure that's consistent
% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
% TODO: Decide on better terms than "fitted" and "not fitted".
% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
% complexity vs. accuracy at the time of convergence.
% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(V^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(V^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(V^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

@@

\subsection*{Analytical models for comparison}
The analytical models used for comparison are also listed in
Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces
added to a collection of aerodynamic terms.  The primary aerodynamic
contribution in all quasi-steady models is a translational lift term
proportional to the squared velocity of each span-element of the wing that is
related to the wing's angle of attack and empirically determined coefficient of
lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal
horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when
out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho V^2AC_L,
\end{equation}
where \(V\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

% TODO: Understand and verify (numerically) how we get an effective v^2 via
% integration along the span, and document that in the code.
